# Bioconductor Package 
## Overview of the BioConductor package 

## Installation of BioConductor package
To install BioConductor run the below command

```
if (!requireNamespace("BiocManager", quitely=TRUE))
     install.packages("BiocManager")
 BiocManager::install("ChemmineR")
```

## Importing Data in R
<at video 1hr 16min>

## Writing Script of Code in R 
Script are useful as they allow one to easily pick up where they left off on a project and progressively build and refine code and analysis. 

They also allow one to easily reproduce analysis run earlier. 

A script is a set of commands that usually includes commenting on what each piece of code intended to do 

## Data Interface
Follow the following steps when working with data from external interfaces; 

- Refer CSv working file(filepath) e.g `test.csv`
- Get the current working directory, `getwd()`
- Set the current working directory `setwd()`. 

<show a csv file screenshot> 

### Reading the *.csv file 
Reading the csv file 

```{r}
proteindata <- read.csv("data/test.csv")
```

and can be converted to a dataframe and explored 

```
# just to confirm if the said data is a data frame 
print(is.data.frame(proteindata))

print(ncol(proteindata)) # count the number of columns 

print(nrow(proteindata)) # count the number of records/rows/observations
```

### Analyzing the *.csv file 

```{r}
csvfile <- read.csv("data/data.csv", header=TRUE)
#csvfile <- read.csv(file.choose(), header=TRUE)
# choose data.csv in the 'data' directory
```

Calculate the mean of the salary
```{r}
# attach the file inorder to avoid the $ sign
attach(csvfile)

mean(Salary)
```

```{r}
# find the column names 
names(csvfile)
```


Now, lets jump to the protein data set 
```{r}
# Load the data set 
proteindata <- read.csv("data/test.csv")
proteindata
```


Convert it to dataframe and count rows, columns 
```{r}
print(is.data.frame(proteindata))
ncol(proteindata)
row(proteindata)
```

Get the protein detail having the most residues and the max residues value.
```{r}
objname1 <- max(proteindata$NO_RESIDUES)
objname1

objname2 <- subset(proteindata, NO_RESIDUES == max(NO_RESIDUES))
objname2
```

Get the protein detail having entry in PDB database
```{r}
objname3 <- subset(proteindata, DBNAME == "PDB")
print(objname3)
```

Get the protein detail having more than 3 dimer from PDB database
```{r}
objname5 <- subset(proteindata, as.Date(DEPOSITED_DATE) > as.Date("2014-05-05"))
print(objname5)
```












