# Bioconductor Package 
## Overview of the BioConductor package 

## Installation of BioConductor package
To install BioConductor run the below command

```
if (!requireNamespace("BiocManager", quitely=TRUE))
     install.packages("BiocManager")
 BiocManager::install("ChemmineR")
```

## Importing Data in R
Depending on the type of data file, R has a varying methods to load the data. The below list shows how to load different types of files to the R program. 

- `data <- read_csv("path/to/your/file.csv")` for csv file
- `data <- read_excel("path/to/your/file.xlsx", sheet = "Sheet1")` for excel file
- `data <- fread("path/to/your/file.csv")` for large csv files.
- `data <- read_sav("path/to/your/file.sav")` for SPSS files.
- `data <- read_dta("path/to/your/file.dta")` for stata files.
- `data <- read_sas("path/to/your/file.sas7bdat")` for SAS files.

## Writing Script of Code in R 
In R, the use can interact directly with the console or prepare a script inform of `*.R` or `*.Rmd` (R Markdown) files. 
Script are useful as they allow one to;

- easily pick up where they left off on a project and progressively build and refine code and analysis. 
- easily reproduce analysis run earlier and documentation. This also saves time on repetetive tasks. 
- do a batch processing such that tasks are automotated instead of running them one by one. 
- debug the code and find errors line by line. 
- version control such as git that tracks the changes and enables collaboration with other developers.

A script is a set of commands that usually includes commenting on what each piece of code intended to do 

## Data Interface
Follow the following steps when working with data from external interfaces; 

- Refer CSV working file(filepath) e.g `test.csv`
- Get the current working directory, `getwd()`
- Set the current working directory `setwd()`. 

### Reading the *.csv file 
In this section, we will work with two csv files, `test.csv` and `data.csv`. Both files are stored in a directory named `data`. Lets load the `data.csv` to the program. 

```{r}
person_info <- read.csv("data/data.csv")
```

The `read.csv()` function automatically converts the loaded to a data frame. `is.data.frame()` is used to find if a variable is a data frame

```{r}
# just to confirm if the said data is a data frame 
print(is.data.frame(person_info))
```
The `person_info` data set is confirmed to be a variable.
The `ncol()` function counts the number of columns/variables/features in a data frame
```{r}
print(ncol(person_info)) # count the number of columns 
```

This data frame has 6 columns. Alternatively, the `nrow()` function counts the number of rows/observations/records. 

```{r}
print(nrow(person_info)) # count the number of records/rows/observations
```

This data set has 10 rows and 6 columns. 

### Brief data analysis
When the function, `attach()` is called to a data frame, the R program can refer to specific column without associating it with the data frame name. In this section the `person_info` will be attached to R and dive shallow into data analysis. 

Calculate the mean of the salary
```{r}
# attach the file inorder to avoid the $ sign
attach(person_info)

mean(Salary)
```

The average salary in the data set is 65200. The `names()` function is called to find the column names of the data set. 
```{r}
# find the column names 
names(person_info)
```


Now, lets jump to the protein data which is stored in the `test.csv` file. The `head()` function is used to show the first 6 records of the dataset
```{r}
# Load the data set 
proteindata <- read.csv("data/test.csv")
head(proteindata)
```


Count the number of rows and columns in the protein dataset
```{r}
ncol(proteindata)
nrow(proteindata)
```

The data frame has 10 rows and 6 columns. Alternatively, the `dim()` can be used, it returns the number of rows and columns respectively. 
```{r}
dim(proteindata)
```

Get the protein detail having the most residues and the max residues value. 
```{r}
objname1 <- max(proteindata$NO_RESIDUES)
objname1

objname2 <- subset(proteindata, NO_RESIDUES == max(NO_RESIDUES))
objname2
```

The `max()` function is used to find the maximum value in a vector or column. In this code, `max(proteindata$NO_RESIDUES)` returns the highest number of residues (`NO_RESIDUES`) found in the `proteindata` dataset, and this value is stored in `objname1`.

The `subset()` function is used to filter rows in a data frame that meet certain conditions. Here, `subset(proteindata, NO_RESIDUES == max(NO_RESIDUES))` returns all rows from the `proteindata` dataset where the `NO_RESIDUES` column matches the maximum value, and the result is stored in `objname2`.

- Get the protein detail having entry in PDB database
```{r}
objname3 <- subset(proteindata, DBNAME == "PDB")
print(objname3)
```

The `proteindata` dataset is filtered to select rows where the `DBNAME` column is equal to `"PDB"`. This means it extracts all proteins that are listed in the PDB database. The resulting subset is stored in `objname3`.


Get the protein detail having more than 3 dimer from PDB database
```{r}
objname5 <- subset(proteindata, as.Date(DEPOSITED_DATE) > as.Date("2014-05-05"))
print(objname5)
```

The subset function is used to filter the `proteindata` dataset to include only rows where `DBNAME` is `"PDB"` and `DIMER_COUNT` is greater than 3.

The filtered data is stored in `objname5`

