# Bioconductor Package 
## Overview of the BioConductor package 

## Installation of BioConductor package
To install BioConductor run the below command

```
if (!requireNamespace("BiocManager", quitely=TRUE))
     install.packages("BiocManager")
 BiocManager::install("ChemmineR")
```

## Importing Data in R
<at video 1hr 16min>

## Writing Script of Code in R 
Script are useful as they allow one to easily pick up where they left off on a project and progressively build and refine code and analysis. 

They also allow one to easily reproduce analysis run earlier. 

A script is a set of commands that usually includes commenting on what each piece of code intended to do 

## Data Interface
Follow the following steps when working with data from external interfaces; 

- Refer CSv working file(filepath) e.g `test.csv`
- Get the current working directory, `getwd()`
- Set the current working directory `setwd()`. 

<show a csv file screenshot> 

### Reading the *.csv file 
Reading the csv file 

```{r}
proteindata <- read.csv("data/test.csv")
```

and can be converted to a dataframe and explored 

```{r}
# just to confirm if the said data is a data frame 
print(is.data.frame(proteindata))

print(ncol(proteindata)) # count the number of columns 

print(nrow(proteindata)) # count the number of records/rows/observations
```

### Analyzing the *.csv file 

Get the max residues from protein data 
```{r}
objname1 <- max(proteindata$NO_RESIDUES)
print(objname1)
```

Get the protein detail having the most residues. 

- `subset()` is used to select and filter variables and observations. 
```{r}
objname2 <- subset(proteindata, NO_RESIDUES == max(NO_RESIDUES))
print(objname2)
```

Get the protein detail having entry in PDB database
```{r}
objname3 <- subset(proteindata, DBNAME == "PDB")
print(objname3)
```

Get the protein detail having more than 3 dimer from PDB database
```{r}
objname4 <- subset(proteindata, NO_DIMER > 3 & DBNAME == "PDB")
print(objname4)
```


